---
title: "Re-evaluation of the Philips curve"
author: "Dominik Gulácsy"
date: "02/01/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
abstract: 'This analysis is addressing the question whether the inverse relationship between unemployment and inflation stated by the Philips curve exists in modern economies. I use cross-sectional data on countries from 2017. After running several regression models, my results support that this inverse relationship does not hold for the present (2016-2018). This finding has relevance to appropiately evaluate policy decisions regarding unemployment and economic growth, and assess the overall health of the economy.'
subtitle: Does higher inflation convey lower unemployment?
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

```{r, results='hide', message=FALSE, warning=FALSE}
# Initialize environment ----------------------------------------------------------------
# Visualization
library(ggthemes)
library(ggplot2)
require(scales)
require(gridExtra)
library(Hmisc)
library(texreg)
# Modeling
library(estimatr)
library(lspline)
# General
library(dplyr)
library(tidyverse)
library(knitr)
# Propitiatory script
source("sum_stat.R", local = knitr::knit_global())

rm(list = ls())
path <- "/Users/Dominik/OneDrive - Central European University/1st_trimester/DA2/DA2_Assignments/Assignment2/"

# Data Prep -------------------------------------------------------------

# Import data
df_all <- read_csv("https://raw.githubusercontent.com/dgulacsy/DA2_Assignments/main/Assignment2/data/clean/unemp_infl_clean.csv")
df_all$year <- as.factor(df_all$year)

# Change unit of measurement for population
df_all$population <- df_all$population/10^5

# Filter for year 2017
df <- filter(df_all, df_all$year==2017)

# Drop observations with an inflation rate below or equal to zero to take reciprocal properly
df <- filter(df, df$inflation > 0)
```

```{r, results='hide', message=FALSE, warning=FALSE}
# Create sum_stat function
sum_stat <- function( df , var_names , stats , num_obs = TRUE ){
  k <- length( var_names )
  built_in_stat <- c('mean','median','mode','min','max','1st_qu.','3rd_qu',
                     'sd','var','range','iqr')
  do_stat <- intersect( stats , built_in_stat )
  if ( is_empty(do_stat) ){
    stop('Error, no such statistics is implemented! Choose from: mean,median,min,max,1st_qu.,3rd_qu')
  }
  # By default add the number of missing observations and the number of used observations
  m <- length( do_stat )
  if ( num_obs ){
    do_stat <- c( do_stat , "# missing", "# used obs")
  }
  # Create tibble for output
  sum_stat <- as_tibble( matrix( 0 , nrow = m , ncol = k ) , name_repair = "unique" )
  for ( j in 1 : k ) {
    # Get the data for the j'th variable
    var_j <- df[ var_names[ j ] ]
    if ( num_obs ){
      # Count the missing values and add to statistics
      sum_stat[ m + 1 , j ] <- as.integer( sum( is.na( var_j ) ) )
      # Count observations used
      sum_stat[ m + 2 , j ] <- as.integer( sum( !is.na( var_j ) ) )
    }
    # Remove missing values
    var_j <- var_j[ !is.na( var_j ) ]
    # Name the sum_stat's column
    colnames( sum_stat )[ j ] <- var_names[ j ]
    for ( i in 1 : m ) {
      # Central tendency
      if (do_stat[ i ] == "mean"){
        sum_stat[[i,j]] <- mean( var_j )
      } else if (do_stat[ i ] == "median"){
        sum_stat[i,j] <- median( var_j )
      } else if (do_stat[ i ] == "mode"){
        sum_stat[i,j] <- mode( var_j )
      } 
      # Support
      else if (do_stat[ i ] == "min"){
        sum_stat[i,j] <- min( var_j )
      } else if (do_stat[ i ] == "max"){
        sum_stat[i,j] <- max( var_j )
      } 
      # Quartiles
      else if (do_stat[ i ] == "1st_qu."){
        sum_stat[i,j] <- quantile( var_j , probs = 0.25 )
      } else if (do_stat[ i ] == "3rd_qu"){
        sum_stat[i,j] <- quantile( var_j , probs = 0.75)
      } 
      # Dispersion
      else if (do_stat[ i ] == "sd"){
        sum_stat[i,j] <- sd( var_j )
      } else if (do_stat[ i ] == "var"){
        sum_stat[i,j] <- var( var_j )
      } else if (do_stat[ i ] == "range"){
        sum_stat[i,j] <- max( var_j ) - min( var_j )
      } else if (do_stat[ i ] == "iqr"){
        sum_stat[i,j] <- quantile( var_j , probs = 0.75) - quantile( var_j , probs = 0.25)
      } 
    }
  }
  # Finally add a column which contains the requested statistics and relocate to first position
  sum_stat <- sum_stat %>% 
    mutate( statistics = do_stat ) %>% 
    relocate( statistics )
  
  return( sum_stat )
}

```
```{r}
# Linear Correlation Coefficient Heatmap ----------------------------------

corrmatrix <- function(df, year) {
  cors <- function(df) {
    M <- rcorr(as.matrix(df)) 
    Mdf <- map(M, ~data.frame(.x)) 
    return(Mdf) }
  
  formatted_cors <- function(df){
    cors(df) %>%
      map(~rownames_to_column(.x, var="measure1")) %>%
      map(~pivot_longer(.x, -measure1, "measure2")) %>% 
      bind_rows(.id = "id") %>%
      pivot_wider(names_from = id, values_from = value) %>%
      mutate(sig_p = ifelse(P < .05, T, F), p_if_sig = ifelse(P <.05, P, NA), r_if_sig = ifelse(P <.05, r, NA)) }
  
  numeric_df <- keep( df , is.numeric)
  
  result <- formatted_cors(numeric_df) %>% 
    ggplot(aes(measure1, measure2, fill=r, label=round(r_if_sig,2))) +
    geom_tile() + 
    labs(x = NULL, y = NULL, fill = "Pearson's\nCorrelation", title=paste0("Linear Correlation Coefficients ",year), subtitle="Only significant coefficients shown") + 
    scale_fill_gradient2(mid="#FBFEF9",low="#0C6291",high="#A63446", limits=c(-1,1)) +
    geom_text() +
    theme_bw() +
    scale_x_discrete(expand=c(0,0)) + 
    scale_y_discrete(expand=c(0,0))

  rm(numeric_df)
  return(result)
}
```

# Introduction
The main goal of this analysis is to test the Philips curve's validity in current times. This economic theory was developed by [A. W. Phillips](https://onlinelibrary.wiley.com/doi/full/10.1111/j.1468-0335.1958.tb00003.x) stating that inflation and unemployment have a stable and inverse relationship. It was based on the idea that with economic growth comes inflation, which in turn should lead to more jobs and less unemployment. It sounds logical, however it was seriously called into question during the times of stagflation in the 1970's.   

The major difference in case of this analysis is that by conducting a cross-sectional causal analysis I can get insight on whether the theory stands in global terms instead of checking it for one country throughout time. To eventually check the theory I run regressions by taking the reciprocal of inflation as this should make the pattern of association linear when there is indeed a an inverse relationship between unemployment and inflation.

# Data
The data used in the analysis is solely from the World Bank. I used the WDI package to download the data in R. The data contains economic indicators on 122 countries for 2017 after removing observations with any missing values. I may add that it is highly probable that by this smaller countries or countries that are in turmoil (e.g.: Venezuela) will be underrepresented as they are more likely to have missing values. Given that there are only few observations available I ignore this issue.  

Otherwise, the quality of the raw data is superior, it is well-documented and provides good coverage. However, one may argue that indicators are not truly comparable and may be politically biased as they are from different national statistical agencies. Since there are no readily available data to counter for these measurement errors I keep in mind this as a limitation to results. Furthermore, to appropriately check the inverse relationship I need to drop observations with inflation below 0 since it messes up the linear regression. This likely makes the sample even less representative as well. Finally, I end up with 118 observations.

As I need to regress unemployment on inflation I want to compare countries that can be considered similar in every other way but inflation. Therefore I use some control variables in my analysis which are the following: GDP Growth (%), Savings (% of GDP), Broad Money (% of GDP) and Government Expenditure (% of GDP).

```{r, fig1, fig.height = 4, fig.width = 9, fig.cap = "Distribution of variables (2017)"}
# EDA ---------------------------------------------------------------------

# Histograms

fig1 <- df %>% rename(
  'Gov. Expenditure (% of GDP)' = govexp,
  'GDP Growth (%)' = gdpgrowth,
  'Broad Money (% of GDP)' = money,
  'Savings (% of GDP)' = savings,
  'Inflation (%)' = inflation,
  'Unemployment (%)' = unemployment) %>% 
  subset (select = -population) %>% 
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
  facet_wrap(~key, scales = "free") +
  geom_histogram()+
  theme_bw() + 
  scale_fill_wsj()

fig1
```

As it can be seen on the histograms all variables are in relative terms. This approach means that size is not relevant. Also, this makes interpretation easier and variables more normal thus improving the fit of the linear regression. Some variables are fairly normal while some are closer to lognormal distribution (Broad Money, Inflation, Unemployment). In case of money supply and unemployment it is likely due to the fact that these measures cannot be negative while in case of inflation it is the result of a technical requirement.  

Taking the logarithm of these variables may alleviate non-normality but interpretation would be hard so I decide not to take logs. Looking at the distributions I can see some extreme values in gov. expenditure and inflation. I investigate their effect as a robustness check later on.

```{r}
# Summary statistics
desc_stat <- sum_stat(df,
         c('govexp','gdpgrowth','money','savings','inflation','unemployment'), 
         c('mean','median','min','max','1st_qu.','3rd_qu','sd','range'),num_obs = F)

# Rename columns and round values
colnames(desc_stat) <- c('statistics','Gov. Expenditure (% of GDP)','GDP Growth (%)','Broad Money (% of GDP)','Savings (% of GDP)','Inflation (%)','Unemployment (%)')
desc_stat[,-1] <-round(desc_stat[,-1],2)

kable(desc_stat, caption = "Descriptive statistics of the variables (2017)")
```

# Model Specifications
```{r}
# Run Regressions ------------------------------------------------------------------
# reg1: No controls, simple linear regression
reg1 <- lm_robust(unemployment ~ inflation, data = df )
#summary( reg1 )

# reg2: No controls, simple linear regression, taking reciprocal of inflation
reg2 <- lm_robust(unemployment ~ I(1/inflation), data = df )
#summary( reg2 )

# Models with control variables
# reg3: control for GDP Growth (P.L.S with knot at 0%)
reg3 <- lm_robust(unemployment ~ inflation + lspline(gdpgrowth , 0), data = df )
#summary( reg3 )

# reg4: reg3 + Government Expenditure (P.L.S with knot at 22%)
reg4 <- lm_robust(unemployment ~ inflation + lspline(gdpgrowth , 0) + lspline(govexp, 22), data = df )
#summary( reg4 )

# reg5: reg4 + Broad Money (P.L.S with knot at 37%)
# Note: This is already 7 variables so additional explanatory variables would not really help
reg5 <- lm_robust(unemployment ~ inflation + lspline(gdpgrowth , 0) + lspline(govexp, 22) + lspline(money, 37), data = df )
#summary( reg5 )

# reg6: reg3 + Broad Money (P.L.S with knot at 37%) + Savings (P.L.S with knot at 15%)
# Note: This is already 7 variables so additional explanatory variables would not really help
reg6 <- lm_robust(unemployment ~ inflation + lspline(gdpgrowth , 0) + lspline(money, 37) + lspline(savings, 15), data = df )
#summary( reg6 )

# reg7: Final model
# Note: This is already 7 variables so additional explanatory variables would not really help
reg7 <- lm_robust(unemployment ~ inflation + lspline(gdpgrowth , 0) + lspline(money, 37), data = df )
#summary( reg7 )

# reg8: Final model with inflation reciprocal
reg8 <- lm_robust(unemployment ~ I(1/inflation) + lspline(gdpgrowth , 0) + lspline(money, 37), data = df )
#summary( reg8 )
```

## 1. Main model



```{r, results = 'asis'}
# HTML --------------------------------------------------------------------
#htmltools::includeHTML(paste0(path,"out/model_comp_finals.html"))

# PDF ---------------------------------------------------------------------
texreg(list(reg7, reg8),
        type = 'pdf',
        custom.header = list("Unemployment rate"=1:2),
        custom.model.names = c("(7)F","(8)F"),
        custom.coef.names = c("intercept","inflation","1/inflation","GDP Growth (<0%)",
                              "GDP Growth (>=0%)", "Broad Money (<37%)", "Broad Money (>=37%)"),

       table = FALSE,
       use.packages = FALSE,
       include.ci = FALSE,
       caption = "Analysis of the relatioship between unemployment and inflation. Data is for 2017 only."
)

```

## 2. Parameter Stability



## 3. Effect of Influential Observations




```{r}



```

# Findings
generalization, external validity
how close it is to causality, 
what other confounders there may be

```{r}



```



# Summary


# Appendix



## I. Model Specification
1. Investigate pattern of association for 2017 (LOESS charts)
```{r, fig17, fig.height = 10, fig.width = 8, fig.align = "center", fig.cap = "Patterns of associations between pairs of variables (2017)"}

plot1 <- ggplot(df , aes(x = inflation, y = unemployment)) +
  geom_point() +
  theme_bw() +
  geom_smooth(method="loess" , formula = y ~ x )+
  labs( subtitle = "Pattern of Association (LOESS)",
        title = "Unemployment and Inflation",
    y = "Unemployment (%)",
    x = "Inflation (%)")

plot2 <- ggplot(df , aes(x = 1/inflation, y = unemployment)) +
  geom_point() +
  theme_bw() +
  geom_smooth(method="loess" , formula = y ~ x )+
  labs( subtitle = "Pattern of Association (LOESS)",
        title = "Unemployment and 1/Inflation",
        y = "Unemployment (%)",
        x = "1/Inflation (%)")

plot3 <- ggplot( df , aes(x = govexp, y = unemployment)) +
  geom_point() +
  theme_bw() +
  geom_smooth(method="loess" , formula = y ~ x )+
  labs(subtitle = "Pattern of Association (LOESS)",
       title = "Unemployment and Gov. Expenditure",
       y = "Unemployment (%)",
       x = "Gov. Expenditure (% of GDP)") 
  
plot4 <- ggplot( df , aes(x = money, y = unemployment)) +
  geom_point() +
  theme_bw() +
  geom_smooth(method="loess" , formula = y ~ x )+
  labs(subtitle = "Pattern of Association (LOESS)",
       title = "Unemployment and Broad Money",
    y = "Unemployment (%)",
    x = "Broad Money (% of GDP)")
  
plot5 <- ggplot( df , aes(x = savings, y = unemployment)) +
  geom_point() +
  theme_bw() +
  geom_smooth(method="loess" , formula = y ~ x )+
  labs(subtitle = "Pattern of Association (LOESS)",
       title = "Unemployment and Savings",
    y = "Unemployment (%)",
    x = "Savings (% of GDP)")
  
plot6 <- ggplot( df , aes(x = gdpgrowth, y = unemployment)) +
  geom_point() +
  theme_bw() +
  geom_smooth(method="loess" , formula = y ~ x )+
  labs(subtitle = "Pattern of Association (LOESS)",
       title = "Unemployment and GDP Growth",
    y = "Unemployment (%)",
    x = "GDP Growth (%)")

fig<-grid.arrange(arrangeGrob(plot1, plot2, plot3, plot4, plot5, plot6, ncol=2, nrow=3), heights=7)

```

```{r fig_corr_17, fig.align = "center", fig.cap = "Corraltion matrix of variables (2017)"}
corrmatrix(df,"2017")
```

2. Model Comparison 2017
```{r}
# Run Regressions ------------------------------------------------------------------

# Avoiding overfitting: ca. 120 observations, 4-6 variables

# Main regression: score4 = b0 + b1*stratio
#   reg1: No controls, simple linear regression
#   reg2: No controls, simple linear regression, taking reciprocal of inflation to check inverse relationship
# Use better reg and control for:
#   reg3: GDP Growth (P.L.S with knot at 0%)
#   reg4: reg3 + Government Expenditure (P.L.S with knot at 22%)
#   reg5: reg4 + Broad Money (P.L.S with knot at 37%)
#   reg6: reg5 + Savings (P.L.S with knot at 15%)

# Note: Weighted regression using population does not make sense variables can only be interpreted on aggregate (country) level

# reg1: No controls, simple linear regression
reg1 <- lm_robust(unemployment ~ inflation, data = df )

# reg2: No controls, simple linear regression, taking reciprocal of inflation
reg2 <- lm_robust(unemployment ~ I(1/inflation), data = df )

# Models with control variables
# reg3: control for GDP Growth (P.L.S with knot at 0%)
reg3 <- lm_robust(unemployment ~ inflation + lspline(gdpgrowth , 0), data = df )

# reg4: reg3 + Government Expenditure (P.L.S with knot at 22%)
reg4 <- lm_robust(unemployment ~ inflation + lspline(gdpgrowth , 0) + lspline(govexp, 22), data = df )

# reg5: reg4 + Broad Money (P.L.S with knot at 37%)
# Note: This is already 7 variables so additional explanatory variables would not really help
reg5 <- lm_robust(unemployment ~ inflation + lspline(gdpgrowth , 0) + lspline(govexp, 22) + lspline(money, 37), data = df )

# reg6: reg3 + Broad Money (P.L.S with knot at 37%) + Savings (P.L.S with knot at 15%)
# Note: This is already 7 variables so additional explanatory variables would not really help
reg6 <- lm_robust(unemployment ~ inflation + lspline(gdpgrowth , 0) + lspline(money, 37) + lspline(savings, 15), data = df )

# reg7: Final model 
# Note: This is already 7 variables so additional explanatory variables would not really help
reg7 <- lm_robust(unemployment ~ inflation + lspline(gdpgrowth , 0) + lspline(money, 37), data = df )

# reg8: Final model with inflation reciprocal
reg8 <- lm_robust(unemployment ~ I(1/inflation) + lspline(gdpgrowth , 0) + lspline(money, 37), data = df )

```

```{r, fig_regs_17, fig.height = 10, fig.width = 8, fig.align = "center", fig.cap = "Modelling unemployment with inflation and 1/inflation (2017)"}

plot1 <- ggplot( data = df, aes( x = inflation, y = unemployment) ) + 
  geom_point( color='black') +
  geom_smooth( formula = y ~ x , method = lm , color = 'red' ) +
  theme_bw() +
  labs(title = "Unemployment and Inflation",
       subtitle = "Reg1: Simple Linear Regression",
       y = "Unemployment (%)",
       x = "Inflation (%)")

plot2 <- ggplot( data = df, aes( x = 1/inflation, y = unemployment) ) + 
  geom_point( color='black') +
  geom_smooth( formula = y ~ I(1/x) , method = lm , color = 'red' ) +
  theme_bw() +
  labs(title = "Unemployment and reciprocal of Inflation",
       subtitle = "Reg2: Simple Linear Regression",
       y = "Unemployment (%)",
       x = "1 / Inflation (%)")

fig<-grid.arrange(arrangeGrob(plot1, plot2, ncol=2, nrow=1))

```

```{r, results = 'asis'}
# HTML --------------------------------------------------------------------
#htmltools::includeHTML(paste0(path,"out/model_comp.html"))

# PDF ---------------------------------------------------------------------
texreg(list(reg1 , reg2 , reg3 , reg4 , reg5, reg6, reg7, reg8),
        type = 'pdf',
        custom.header = list("Unemployment rate"=1:8),
        custom.model.names = c("(1)","(2)","(3)","(4)","(5)","(6)","(7)F","(8)F"),
        custom.coef.names = c("intercept","inflation","1/inflation","GDP Growth (<0%)","GDP Growth (>=0%)",
                              "Gov. Exp. (<22%)","Gov. Exp. (>=22%)", "Broad Money (<37%)", "Broad Money (>=37%)",
                              "Savings (<15%)", "Savings (>=15%)"),

       table = FALSE,
       use.packages = FALSE,
       include.ci = FALSE,
       caption = "Analysis of the relatioship between unemployment and inflation. Data is for 2017 only."
)
```

## II. Robustness Check
1. Dropping observations that have government expenditure above 30%
```{r}
# HTML --------------------------------------------------------------------
# htmltools::includeHTML(paste0(path,"out/model_comp_case1.html"))

# PDF ---------------------------------------------------------------------

```

2. Dropping observations that have inflation above 30%
```{r}
# HTML --------------------------------------------------------------------
# htmltools::includeHTML(paste0(path,"out/model_comp_case2.html"))

# PDF ---------------------------------------------------------------------

```

## III. Checking External Validity

1. Investigate pattern of association for 2016 (LOESS charts)
```{r, fig16, fig.height = 10, fig.width = 8, fig.align = "center"}

# Filter for year 2016
df <- filter(df_all, df_all$year==2016)

# Drop observations with an inflation rate below or equal to zero to take reciprocal properly
df <- filter(df, df$inflation > 0)

# Drop influential observations regarding government exp
df <- filter(df, df$govexp < 30)

plot1 <- ggplot(df , aes(x = inflation, y = unemployment)) +
  geom_point() +
  theme_bw() +
  geom_smooth(method="loess" , formula = y ~ x )+
  labs( subtitle = "Pattern of Association (LOESS)",
        title = "Unemployment and Inflation",
    y = "Unemployment (%)",
    x = "Inflation (%)")

plot2 <- ggplot(df , aes(x = 1/inflation, y = unemployment)) +
  geom_point() +
  theme_bw() +
  geom_smooth(method="loess" , formula = y ~ x )+
  labs( subtitle = "Pattern of Association (LOESS)",
        title = "Unemployment and 1/Inflation",
        y = "Unemployment (%)",
        x = "1/Inflation (%)")

plot3 <- ggplot( df , aes(x = govexp, y = unemployment)) +
  geom_point() +
  theme_bw() +
  geom_smooth(method="loess" , formula = y ~ x )+
  labs(subtitle = "Pattern of Association (LOESS)",
       title = "Unemployment and Gov. Expenditure",
       y = "Unemployment (%)",
       x = "Gov. Expenditure (% of GDP)") 
  
plot4 <- ggplot( df , aes(x = money, y = unemployment)) +
  geom_point() +
  theme_bw() +
  geom_smooth(method="loess" , formula = y ~ x )+
  labs(subtitle = "Pattern of Association (LOESS)",
       title = "Unemployment and Broad Money",
    y = "Unemployment (%)",
    x = "Broad Money (% of GDP)")
  
plot5 <- ggplot( df , aes(x = savings, y = unemployment)) +
  geom_point() +
  theme_bw() +
  geom_smooth(method="loess" , formula = y ~ x )+
  labs(subtitle = "Pattern of Association (LOESS)",
       title = "Unemployment and Savings",
    y = "Unemployment (%)",
    x = "Savings (% of GDP)")
  
plot6 <- ggplot( df , aes(x = gdpgrowth, y = unemployment)) +
  geom_point() +
  theme_bw() +
  geom_smooth(method="loess" , formula = y ~ x )+
  labs(subtitle = "Pattern of Association (LOESS)",
       title = "Unemployment and GDP Growth",
    y = "Unemployment (%)",
    x = "GDP Growth (%)")

fig<-grid.arrange(arrangeGrob(plot1, plot2, plot3, plot4, plot5, plot6, ncol=2, nrow=3), heights=7)

```

2. Model Comparison 2016

```{r}
# HTML --------------------------------------------------------------------
#htmltools::includeHTML(paste0(path,"out/model_comp_2016.html"))

# PDF ---------------------------------------------------------------------

```

3. Investigate pattern of association for 2018 (LOESS charts)

```{r, fig18, fig.height = 10, fig.width = 8, fig.align = "center"}

# Filter for year 2018
df <- filter(df_all, df_all$year==2018)

# Drop observations with an inflation rate below or equal to zero to take reciprocal properly
df <- filter(df, df$inflation > 0)

# Drop influential observations regarding government exp
df <- filter(df, df$govexp < 30)

# Drop influential observations regarding savings
df <- filter(df, df$savings > 0)

# Investigation of association patterns, plot loess for pairs of variables
plot1 <- ggplot(df , aes(x = inflation, y = unemployment)) +
  geom_point() +
  theme_bw() +
  geom_smooth(method="loess" , formula = y ~ x )+
  labs( subtitle = "Pattern of Association (LOESS)",
        title = "Unemployment and Inflation",
    y = "Unemployment (%)",
    x = "Inflation (%)")

plot2 <- ggplot(df , aes(x = 1/inflation, y = unemployment)) +
  geom_point() +
  theme_bw() +
  geom_smooth(method="loess" , formula = y ~ x )+
  labs( subtitle = "Pattern of Association (LOESS)",
        title = "Unemployment and 1/Inflation",
        y = "Unemployment (%)",
        x = "1/Inflation (%)")

plot3 <- ggplot( df , aes(x = govexp, y = unemployment)) +
  geom_point() +
  theme_bw() +
  geom_smooth(method="loess" , formula = y ~ x )+
  labs(subtitle = "Pattern of Association (LOESS)",
       title = "Unemployment and Gov. Expenditure",
       y = "Unemployment (%)",
       x = "Gov. Expenditure (% of GDP)") 
  
plot4 <- ggplot( df , aes(x = money, y = unemployment)) +
  geom_point() +
  theme_bw() +
  geom_smooth(method="loess" , formula = y ~ x )+
  labs(subtitle = "Pattern of Association (LOESS)",
       title = "Unemployment and Broad Money",
    y = "Unemployment (%)",
    x = "Broad Money (% of GDP)")
  
plot5 <- ggplot( df , aes(x = savings, y = unemployment)) +
  geom_point() +
  theme_bw() +
  geom_smooth(method="loess" , formula = y ~ x )+
  labs(subtitle = "Pattern of Association (LOESS)",
       title = "Unemployment and Savings",
    y = "Unemployment (%)",
    x = "Savings (% of GDP)")
  
plot6 <- ggplot( df , aes(x = gdpgrowth, y = unemployment)) +
  geom_point() +
  theme_bw() +
  geom_smooth(method="loess" , formula = y ~ x )+
  labs(subtitle = "Pattern of Association (LOESS)",
       title = "Unemployment and GDP Growth",
    y = "Unemployment (%)",
    x = "GDP Growth (%)")

fig<-grid.arrange(arrangeGrob(plot1, plot2, plot3, plot4, plot5, plot6, ncol=2, nrow=3), heights=7)

```

4.Model Comparison 2018

```{r}
# HTML --------------------------------------------------------------------
#htmltools::includeHTML(paste0(path,"out/model_comp_2018.html"))

# PDF ---------------------------------------------------------------------

```
